---
title: "01_data_processing"
author: "Flora"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

# loading packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(semTools))

options(scipen = 999)
```


# importing dataset
```{r message=FALSE, warning=FALSE}
dat_cleaned <- read_csv("data/data_cleaned.csv")
```


# tidying data

## reordering and renaming variables
```{r warning=FALSE}
# reordering variables
dat_cleaned <- dat_cleaned %>% 
  select(id = id_emo, day, beep, age, gender, edu, date_emotics, ended_pre, date_diary, diary_counter,
         sad:subc_psycho, cheerful = cheerfull, irritated:event_pleasantness, #esm
         cerq1:cerq36, sb_cerq:ob_cerq) # cross-sectional


# renaming variables for consistency
dat_cleaned <- dat_cleaned %>% 
  rename_at(vars(starts_with("cerq")), funs(str_replace(., "cerq", "cerq_")))
```


## filling missing values in cross-sectional variables
```{r}
dat_cleaned <- dat_cleaned %>%
  group_by(id) %>% 
  fill(c(sb_cerq:ob_cerq), .direction = "updown") %>% 
  ungroup()
```


## filtering obeservations with missing ESM data
```{r}
dat_cleaned <- dat_cleaned %>% 
  filter(!is.na(date_emotics) & !is.na(day))
```


# CFA and reliability

## CFA for rumination
```{r}
dat_cfa_rum <- 
  dat_cleaned %>% 
  dplyr::select(id, cerq_1, cerq_9, cerq_19, cerq_29) %>%
  distinct()
  
cfa_rum <- 'trait_rumination =~ cerq_1 + cerq_9 + cerq_19 + cerq_29'

fit_rumi <- cfa(model = cfa_rum, data = dat_cfa_rum, std.lv = TRUE, estimator = "ML")
```

```{r}
fitMeasures(fit_rumi, c("chisq", "df", "pvalue", "cfi",
                        "tli", "rmsea", "srmr"))
```

```{r}
summary(fit_rumi, fit.measures = TRUE, standardize = TRUE)
```

```{r}
# reliability of rumination factor (using omega)
(reliability(fit_rumi))
```
```{r}
# factor loadings
inspect(fit_rumi, what="std")$lambda
```

```{r}
# adding rumination between-subject factor scores to main dataset
scores_cfa_rumination <-
  dat_cfa_rum %>%
  distinct() %>%
  cbind(lavPredict(fit_rumi, newdata = dat_cfa_rum[1:5])) %>%
  select(id, trait_rumination)

dat_cleaned <-
  dat_cleaned %>%
  left_join(scores_cfa_rumination, by = "id")
```


# multilevel CFA for negative affective states
```{r, warning = F}
cfa_aff <-
  dat_cleaned %>%
  filter(!is.na(date_emotics)) %>%
  dplyr::select(id, date_emotics, beep, sad, afraid, unrested, irritated, angry) %>%
  filter_at(vars(sad:angry), all_vars(!is.na(.))) %>%
  distinct() %>%
  ungroup()

cfa_aff_syntax <- '
  level: 1
      NA_within =~ sad + afraid + unrested + angry
      afraid ~~ angry

  level: 2
      NA_between =~ sad + afraid + unrested + angry
'
# "irritated" was omitted from the analysis as it worsened model fit

fit_aff <- cfa(model = cfa_aff_syntax, data = cfa_aff, cluster = "id", std.lv = TRUE, estimator = "ML")
```

```{r}
fitMeasures(fit_aff, c("chisq", "df", "pvalue", "cfi", "tli",
                       "rmsea", "srmr_within", "srmr_between"))
```

```{r}
summary(fit_aff, fit.measures = TRUE, standardize = TRUE)
```

```{r}
modificationindices(fit_aff) %>%
  dplyr::select("lhs", "op", "rhs", "level", "mi") %>%
  arrange(desc(mi)) %>%
  # filter(level == 1) %>% 
  head(n = 10) %>%
  print()
```

```{r}
# within-person factor loadings
(inspect(fit_aff, what = "std")$within$lambda)

# between-person factor loadings
(inspect(fit_aff, what = "std")$id$lambda)

# reliability (using omega)
(reliability(fit_aff))
```

```{r}
# adding between subject factor scores to within subject factor scores
dat_cfa_lev_2 <-
  cfa_aff %>%
  dplyr::select(id) %>%
  distinct() %>%
  cbind(lavPredict(fit_aff, newdata = cfa_aff[4:8], level = 2L))

dat_cfa_lev_1 <-
  cfa_aff %>%
  cbind(lavPredict(fit_aff, newdata = cfa_aff[4:8], level = 1L)) %>%
  dplyr::select(-c(sad:angry))

dat_cfa <-
  dat_cfa_lev_1 %>%
  left_join(dat_cfa_lev_2, by = "id") %>%
  mutate(NA_sum =  NA_within + NA_between)

dat_cleaned <-
  dat_cleaned %>%
  left_join(dat_cfa, by = c("id", "date_emotics", "beep"))
```


# computing and centering aggregated mean of state rumintaion and event unpleasantness
```{r}
# rumination
dat_cleaned <- dat_cleaned %>% 
  group_by(id) %>% 
  mutate(mean_state_rumi = mean(regu_rumination, na.rm = TRUE), # between-person means
         rumination_centered = scale(regu_rumination, center = T, scale = F) # within-person centering
         ) %>% 
  ungroup()

dat_cleaned <- dat_cleaned %>% 
  mutate(mean_state_rumi_grand_centered = scale(mean_state_rumi, center = T, scale = F)) # grand-mean centering


# event unpleasantness
  ## reversing event pleasantness to event unpleasantness
  dat_cleaned <-
    dat_cleaned %>% 
    mutate(event_unpleasantness = 8-event_pleasantness) %>% 
    dplyr::select(-event_pleasantness)
  
  
  ## within-person centering of event unpleasantness and calculating between-person means
  dat_cleaned <-
    dat_cleaned %>% 
    group_by(id) %>%
    mutate(event_unpleasantness_centered = scale(event_unpleasantness, center = T, scale = F), # within-person centering
           event_unpleasantness_mean = mean(event_unpleasantness, na.rm = T) # between-person means
           ) %>%
    distinct()
```


# creating .csv file for running descriptive statistics
```{r}
dat_cleaned %>% 
  write.csv("data/data_processed.csv", row.names = F)
```


# log10-transforming NA_sum
```{r}
dat_cleaned <- dat_cleaned %>% 
  mutate(NA_sum_log = log10(NA_sum + 2))
```


# selecting model-relevant variables and converting gender into factor
```{r}
dat_model <- dat_cleaned %>%
  select(id:edu, event_unpleasantness_centered, trait_rumination, NA_within:NA_sum, NA_sum_log, mean_state_rumi, rumination_centered, mean_state_rumi_grand_centered) %>% 
  mutate(gender = factor(gender, levels = c(1, 2, 3), labels = c("Female", "Male", "Other"))
  )
```


# distribution of NA_sum
```{r}
hist(dat_model$NA_sum)
hist(dat_model$NA_sum_log)

```


# writing dat_model into .csv file for running models
```{r}
dat_model %>%
  write.csv("data/data_model.csv", row.names = F)
```

