---
title: "01_data_processing"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# loading packages
```{r}

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(sjPlot))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(lavaan))
suppressPackageStartupMessages(library(semTools))
suppressPackageStartupMessages(library(semPlot))
# suppressPackageStartupMessages(library(ltm))

options(scipen = 999)
```


# reading dataset
```{r message=FALSE, warning=FALSE}
dat_cleaned <- read_csv("data/data_cleaned.csv")
```


# reordering and renaming variables
```{r warning=FALSE}
# reordering variables
dat_cleaned <- dat_cleaned %>% 
  select(id = id_emo, day, beep, age, gender, edu, date_emotics, ended_pre, date_diary, diary_counter,
         sad:subc_psycho, cheerful = cheerfull, irritated:event_pleasantness, #esm
         cerq1:cerq36, sb_cerq:ob_cerq) # cross-sectional

# renaming variables for consistency
dat_cleaned <- dat_cleaned %>% 
  rename_at(vars(starts_with("cerq")), funs(str_replace(., "cerq", "cerq_")))
```


# filling missing values in cross-sectional variables
```{r}
dat_cleaned <- dat_cleaned %>%
  group_by(id) %>% 
  fill(c(sb_cerq:ob_cerq), .direction = "updown") %>% 
  ungroup()
```


# filtering obeservations with missing ESM
```{r}
dat_cleaned <- dat_cleaned %>% 
  filter(!is.na(date_emotics) & !is.na(day))
```


# parallel analysis of affective states
```{r}
# prin <-
#   dat_cleaned %>%
#   select(sad, calm, afraid, unrested, enthusiastic, angry, worried, cheerful, irritated, active)
# 
# (parallel <- fa.parallel(prin, fm = "pa"))
# 
# (principal <-
#     psych::principal(prin, nfactors = 2, rotate = "oblimin"))
# 
# # binding component scores to dataset
# dat_cleaned <-
#   dat_cleaned %>%
#   cbind(as.data.frame(principal$scores)) %>%
#   mutate(negaff = TC1,
#          posaff = TC2) %>%
#   select(-TC1, -TC2)
```


# CFA for rumination
```{r}
dat_cfa_rum <- 
  dat_cleaned %>% 
  dplyr::select(id, cerq_1, cerq_9, cerq_19, cerq_29) %>%
  distinct()
  
cfa_rum <- 'trait_rumination =~ cerq_1 + cerq_9 + cerq_19 + cerq_29'

fit_rumi <- cfa(model = cfa_rum, data = dat_cfa_rum, std.lv = TRUE, estimator = "ML")
```

```{r}
fitMeasures(fit_rumi, c("chisq", "df", "pvalue", "cfi", "tli",
                   "rmsea", "srmr"))
```

```{r}
summary(fit_rumi, fit.measures = TRUE, standardize = TRUE)
```

```{r}
# reliability of rumination factor (omega)
(reliability(fit_rumi))
```
```{r}
# factor loadings
inspect(fit_rumi,what="std")$lambda
```



```{r}
# add rumination between-subject factor scores to the main dataset
scores_cfa_rumination <-
  dat_cfa_rum %>%
  distinct() %>%
  cbind(lavPredict(fit_rumi, newdata = dat_cfa_rum[1:5])) %>%
  select(id, trait_rumination)

dat_cleaned <-
  dat_cleaned %>%
  left_join(scores_cfa_rumination, by = c("id"))
```


# multilevel CFA for negative affective states
```{r, warning = F}
cfa_aff <-
  dat_cleaned %>%
  filter(!is.na(date_emotics)) %>%
  dplyr::select(id, date_emotics, beep, sad, afraid, unrested, irritated, angry) %>%
  filter_at(vars(sad:angry), all_vars(!is.na(.))) %>%
  distinct() %>%
  ungroup()

cfa_aff_syntax <- '

  level: 1

      NA_within =~ sad + afraid + unrested + angry
      afraid ~~ angry

  level: 2

      NA_between =~ sad + afraid + unrested + angry
'
# "irritated" has been omitted from the analysis since it worsened the fit of the model

fit <- cfa(model = cfa_aff_syntax, data = cfa_aff, cluster = "id", std.lv = TRUE, estimator = "ML")
```

```{r}
fitMeasures(fit, c("chisq", "df", "pvalue", "cfi", "tli",
                   "rmsea", "srmr_within", "srmr_between"))
```

```{r}
summary(fit, fit.measures = TRUE, standardize = TRUE)
```

```{r}
modificationindices(fit) %>%
  dplyr::select("lhs", "op", "rhs", "level", "mi") %>%
  arrange(desc(mi)) %>%
  # filter(level == 1) %>% 
  head(n = 10) %>%
  print()
```

```{r}
# within-person factor loadings
(inspect(fit, what = "std")$within$lambda)

# between-person factor loadings
(inspect(fit, what = "std")$id$lambda)

# reliability of NA model
(reliability(fit))
```

```{r}
# between subject factor score-okat hozzáadni a within subject-ekhez
dat_cfa_lev_2 <-
  cfa_aff %>%
  dplyr::select(id) %>%
  distinct() %>%
  cbind(lavPredict(fit, newdata = cfa_aff[4:8], level = 2L))

dat_cfa_lev_1 <-
  cfa_aff %>%
  cbind(lavPredict(fit, newdata = cfa_aff[4:8], level = 1L)) %>%
  dplyr::select(-c(sad:angry))

dat_cfa <-
  dat_cfa_lev_1 %>%
  left_join(dat_cfa_lev_2, by = "id") %>%
  mutate(NA_sum =  NA_within + NA_between)

dat_cleaned <-
  dat_cleaned %>%
  left_join(dat_cfa, by = c("id", "date_emotics", "beep"))
```


# reliability of CERQ rumination scale
```{r}
# dat_cleaned_cerq <- dat_cleaned %>%  
#   dplyr::select(cerq_1, cerq_9, cerq_19, cerq_29) %>% 
#   distinct()
# 
# cronbach.alpha(dat_cleaned_cerq)
```


# computing and centering aggregated mean of state rumintaion and event unpleasantness
```{r}
# rumination
dat_cleaned <- dat_cleaned %>% 
  group_by(id) %>% 
  mutate(mean_state_rumi = mean(regu_rumination, na.rm = TRUE), # between-person means
         rumination_centered = scale(regu_rumination, center = T, scale = F) # within-person centering
         ) %>% 
  ungroup()

dat_cleaned <- dat_cleaned %>% 
  mutate(mean_state_rumi_grand_centered = scale(mean_state_rumi, center = T, scale = F)) # grand-mean centering


# event unpleasantness
  # reversing event pleasantness to event unpleasantness
  dat_cleaned <-
    dat_cleaned %>% 
    mutate(event_unpleasantness = 8-event_pleasantness) %>% 
    dplyr::select(-event_pleasantness)
  
  
  # within-person centering of event unpleasantness and calculating between-person means
  dat_cleaned <-
    dat_cleaned %>% 
    group_by(id) %>%
    mutate(event_unpleasantness_centered = scale(event_unpleasantness, center = T, scale = F), # within-person centering
           event_unpleasantness_mean = mean(event_unpleasantness, na.rm = T) # between-person means
           ) %>%
    distinct()


# creating .csv file
dat_cleaned %>% 
  write.csv("data/data_processed.csv", row.names = F)
```


# log10-transforming NA_sum
```{r}
dat_cleaned <- dat_cleaned %>% 
  mutate(NA_sum_log = log10(NA_sum + 2))
```


# selecting model-relevant variables and converting gender into factor
```{r}
dat_model <- dat_cleaned %>%
  select(id:edu, event_unpleasantness_centered, trait_rumination, NA_within:NA_sum, NA_sum_log, mean_state_rumi, rumination_centered, mean_state_rumi_grand_centered) %>% 
  mutate(gender = factor(gender, levels = c(1, 2, 3), labels = c("Female", "Male", "Other"))
  )
```


# distribution of NA_sum
```{r}
hist(dat_model$NA_sum)
hist(dat_model$NA_sum_log)
```


# writing dat_model into .csv file
```{r}
dat_model %>%
  write.csv("data/data_model.csv", row.names = F)
```

