---
title: "04_descriptive_stats"
author: "Flora"
date: "2022-08-25"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# loading packages

```{r}
options(scipen = 999)

suppressPackageStartupMessages(library(sjPlot))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(confintr))
suppressPackageStartupMessages(library(ppcor))
suppressPackageStartupMessages(library(rcompanion))
suppressPackageStartupMessages(library(showtext))
suppressPackageStartupMessages(library(extrafont))
suppressPackageStartupMessages(library(tidyverse))

loadfonts(device = "all")
font_add(family = "Cooper Hewitt", regular = "CooperHewitt-Medium.otf")
showtext_auto()
```

# importing datasets

```{r message=FALSE}
dat_pre <- read_xlsx("data/dat_pre.xlsx")
dat_processed <- read_csv("data/data_processed.csv")
dat_unfiltered <- read_csv("data/data_unfiltered.csv")
```

# filtering datasets

```{r}
# cleaning pre
dat_pre_clean <- dat_pre %>% 
  mutate(ended_pre = as_datetime(ended)) %>% 
  select(session, ended_pre, age, gender, edu, cerq1, cerq9, cerq19, cerq29) %>% 
  filter(!is.na(ended_pre), age != "under_18") %>% 
  mutate(age = as.numeric(age),
         rum_cerq = rowSums(select(., c("cerq1", "cerq9", "cerq19", "cerq29")))
  ) %>% 
  select(-cerq1, -cerq9, -cerq19, -cerq29)


# filtering sample not included
dat_processed_distinct <- dat_processed %>% 
  select(id, ended_pre, age, gender, edu, rum_cerq) %>% 
  distinct()

dat_not_included <- dat_pre_clean %>% 
  anti_join(dat_processed_distinct, by = c("age", "gender", "edu", "rum_cerq", "ended_pre"))
```

# sample demographics

## cross-sectional = initial

```{r}
# number and age of participants  
dat_pre_clean %>% 
  select(session, age) %>% 
  summarise(N         = n(),
            Mean      = round(mean(age), digits = 2),
            Median    = round(median(age), digits = 2),
            SD        = round(sd(age), digits = 2),
            Range_min = min(age),
            Range_max = max(age))

# gender
dat_pre_clean %>% 
  select(gender) %>% 
  table()
  
dat_pre_clean %>%
  select(gender) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)
  
  # 1 = Female
  # 2 = Male
  # 3 = Other
  

# education
dat_pre_clean %>% 
  select(edu) %>% 
  table()
  
dat_pre_clean %>% 
  select(edu) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

  # 1 = Primary school or lower
  # 3 = Vocational school without high school diploma
  # 4 = High school diploma or equivalent 
  # 5 = Bachelor's or Master's degree
  # 6 = Doctorate (PhD)
  # 7 = Other

# trait rumination
dat_pre_clean %>%
  summarise(N         = n(),
            Mean      = round(mean(rum_cerq), digits = 2),
            Median    = round(median(rum_cerq), digits = 2),
            SD        = round(sd(rum_cerq), digits = 2),
            Range_min = min(rum_cerq),
            Range_max = max(rum_cerq))
```

## not included

```{r}
# number and age of participants  
dat_not_included %>% 
  select(session, age) %>% 
  summarise(N         = n(),
            Mean      = round(mean(age), digits = 2),
            Median    = round(median(age), digits = 2),
            SD        = round(sd(age), digits = 2),
            Range_min = min(age),
            Range_max = max(age))

# gender
dat_not_included %>% 
  select(gender) %>% 
  table()
  
dat_not_included %>%
  select(gender) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)
  
  # 1 = Female
  # 2 = Male
  # 3 = Other
  

# education
dat_not_included %>% 
  select(edu) %>% 
  table()
  
dat_not_included %>% 
  select(edu) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

  # 1 = Primary school or lower
  # 3 = Vocational school without high school diploma
  # 4 = High school diploma or equivalent 
  # 5 = Bachelor's or Master's degree
  # 6 = Doctorate (PhD)
  # 7 = Other

# trait rumination
dat_not_included %>%
  summarise(N         = n(),
            Mean      = round(mean(rum_cerq), digits = 2),
            Median    = round(median(rum_cerq), digits = 2),
            SD        = round(sd(rum_cerq), digits = 2),
            Range_min = min(rum_cerq),
            Range_max = max(rum_cerq))
```

## proceeded to ESM phase = final sample

```{r}
# number and age of participants  
dat_processed %>%
    select(id, age) %>%
    distinct() %>%
    summarise(N         = n(),
              Mean      = round(mean(age), digits = 2),
              Median    = round(median(age), digits = 2),
              SD        = round(sd(age), digits = 2),
              Range_min = min(age),
              Range_max = max(age))


# gender
dat_processed %>% 
  select(id, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table()

dat_processed %>% 
  select(id, gender) %>%
  distinct() %>% 
  select(gender) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)

# 1 = Female
  # 2 = Male
  # 3 = Other

  
# education
dat_processed %>% 
  select(id, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table()

dat_processed %>% 
  select(id, edu) %>%
  distinct() %>% 
  select(edu) %>% 
  table() %>% 
  prop.table() %>% `*`(100) %>% round(2)
  
  # 1 = Primary school or lower
  # 3 = Vocational school without high school diploma
  # 4 = High school diploma or equivalent 
  # 5 = Bachelor's or Master's degree
  # 6 = Doctorate (PhD)
  # 7 = Other


# trait rumination
dat_processed %>%
  select(id, age, gender, edu, rum_cerq) %>% 
  distinct() %>% 
  summarise(N         = n(),
            Mean      = round(mean(rum_cerq), digits = 2),
            Median    = round(median(rum_cerq), digits = 2),
            SD        = round(sd(rum_cerq), digits = 2),
            Range_min = min(rum_cerq),
            Range_max = max(rum_cerq))
```

# descriptives

## descriptives of ESM surveys in final sample

```{r}
# total number of completed esm surveys
dat_processed %>%
  distinct() %>% 
  summarise(n_id  = length(unique(id)),
            n_obs = length(sad))


#...and per capita
dat_processed %>%
  distinct() %>% 
  group_by(id) %>% 
  summarise(n_obs = length(date_emotics)) %>% 
  ungroup() %>% 
  summarise(max_n_obs    = max(n_obs),
            min_n_obs    = min(n_obs),
            mean_n_obs   = mean(n_obs),
            median_n_obs = median(n_obs),
            sd_n_obs     = sd(n_obs))


# descriptives of rumination 
  # state rumination
  dat_processed %>% 
    select(id, day, beep, regu_rumination) %>%
    summarise(
      max_state_rumi    = max(regu_rumination),
      min_state_rumi    = min(regu_rumination),
      mean_state_rumi   = mean(regu_rumination),
      median_state_rumi = median(regu_rumination),
      sd_state_rumi     = sd(regu_rumination)
  )
  
  # mean state rumination
  dat_processed %>% 
    select(id, day, beep, mean_state_rumi) %>%
    summarise(
      max_mean_state_rumi    = max(mean_state_rumi),
      min_mean_state_rumi    = min(mean_state_rumi),
      mean_mean_state_rumi   = mean(mean_state_rumi),
      median_mean_state_rumi = median(mean_state_rumi),
      sd_mean_state_rumi     = sd(mean_state_rumi)
  )


# descriptives of negative affect
dat_processed %>%
  select(id, day, beep, sad, afraid, unrested, angry) %>% 
  mutate(negaff_raw_sum = rowSums(select(., c("sad", "afraid", "unrested", "angry")))) %>% 
  summarise(
    max_negaff    = max(negaff_raw_sum),
    min_negaff    = min(negaff_raw_sum),
    mean_negaff   = mean(negaff_raw_sum),
    median_negaff = median(negaff_raw_sum),
    sd_negaff     = sd(negaff_raw_sum)
  )


# descriptives of event pleasantness
dat_processed %>% 
  select(id, day, beep, event_unpleasantness) %>% 
  summarise(
    min_evunpleasant    = min(event_unpleasantness),
    max_evunpleasant    = max(event_unpleasantness),
    mean_evunpleasant   = mean(event_unpleasantness),
    median_evunpleasant = median(event_unpleasantness),
    sd_evunpleasant     = sd(event_unpleasantness)
  )
```

## statistical tests to compare initial and final samples

```{r}
# test for median age
  # normality test
  shapiro.test(dat_not_included$age)
  shapiro.test(dat_processed_distinct$age)
    # -> normality not assumed -> mann-whitney
  
  # mann-whitney test
  wilcox.test(dat_not_included$age, dat_processed_distinct$age)
  

# test for gender distribution
  gender_compare_not_included <- tibble(gender = dat_not_included$gender, group = "not_included")
  gender_compare_final <- tibble(gender = dat_processed_distinct$gender, group = "final")
  gender_compare <- gender_compare_not_included %>% bind_rows(gender_compare_final)
  
  chisq.test(gender_compare$group, gender_compare$gender)


# test for education
  edu_compare_not_included <- tibble(edu = dat_not_included$edu, group = "not_included")
  edu_compare_final <- tibble(edu = dat_processed_distinct$edu, group = "final")
  edu_compare <- edu_compare_not_included %>% bind_rows(edu_compare_final)
  
  edu_chisq <- chisq.test(edu_compare$group, edu_compare$edu)
  cramersv(edu_chisq)
  

# test for median mss scores
  # normality test
  shapiro.test(dat_not_included$rum_cerq)
  shapiro.test(dat_processed_distinct$rum_cerq)
    # -> normality not assumed -> mann-whitney
  
  # mann-whitney test
  wilcox.test(dat_not_included$rum_cerq, dat_processed_distinct$rum_cerq)
```

# calculating compliance in final sample

## compliance on ESM surveys

```{r}
# filter dataset
dat_unfiltered <- dat_unfiltered %>% 
  filter(!is.na(day) & day != 0)


# identify day of quitting
dat_maxday_esm <- dat_unfiltered %>% 
  select(id_emo, date_emotics, day) %>% 
  filter(!is.na(date_emotics)) %>% 
  group_by(id_emo) %>% 
  summarise(max_day_esm = max(day))

dat_unfiltered <- dat_unfiltered %>% 
  left_join(dat_maxday_esm, by = "id_emo")


# calculating compliance
dat_unfiltered %>%
  select(id_emo, date_emotics, day, beep, max_day_esm) %>%
  group_by(id_emo) %>%
  summarise(completed = sum(!is.na(date_emotics)),
            all = max_day_esm * 8, 
            esm_compl_final = (completed / all) * 100
  ) %>% 
  ungroup() %>%
  summarise(median_final_compl = median(esm_compl_final, na.rm = T),
            mean_final_compl   = mean(esm_compl_final, na.rm = T),
            sd_final_compl     = sd(esm_compl_final, na.rm = T),
            min_final_compl    = min(esm_compl_final, na.rm = T),
            max_final_compl    = max(esm_compl_final, na.rm = T)
  )

# distribution of the number of completed surveys
dat_unfiltered_figure <- dat_unfiltered %>% 
  select(id_emo, date_emotics, day, beep, max_day_esm) %>%
  group_by(id_emo) %>%
  summarise(completed = sum(!is.na(date_emotics)),
            all = max_day_esm * 8, 
            esm_compl_final = (completed / all) * 100
  ) %>% 
  ungroup() %>% 
  distinct()
  
ggplot(dat_unfiltered_figure) +
  geom_histogram(aes(esm_compl_final), bins = 20, fill = "#A9A9A9") +
  theme_classic() +
  labs(
    x = "Compliance rate (%)",
    y = "Number of participants"
  ) +
  theme(axis.title = element_text(size = 12))
```

# correlation between trait and state rumination measures

## correlation

```{r}
corr_rumi <- dat_processed %>% 
  select(id, mean_state_rumi_grand_centered, trait_rumination) %>% 
  distinct() %>% 
  select(-id)

shapiro.test(corr_rumi$trait_rumination)
shapiro.test(corr_rumi$mean_state_rumi_grand_centered)
# normality not assumed

# corr_coeff <- cor.test(corr_rumi$trait_rumination, corr_rumi$mean_state_rumi, method = "spearman")

ci_cor(corr_rumi, probs = c(0.025, 0.975), method = "spearman", type = "bootstrap")

corr_rumi_z <- dat_processed %>% 
  mutate(mean_state_rumi_z = scale(mean_state_rumi),
         rum_cerq_z = scale(trait_rumination)
         ) %>% 
  select(id, mean_state_rumi_z, rum_cerq_z) %>% 
  distinct() %>% 
  select(-id)
```

## partial correlation

```{r}
# filter and join
dat_unfiltered_maxday <- dat_unfiltered %>% 
  select(id_emo, max_day_esm) %>% 
  distinct()

dat_processed <- dat_processed %>% 
  left_join(
    dat_unfiltered_maxday, by = join_by(id == id_emo)
            )

corr_rumi_partial <- dat_processed %>% 
  select(id, mean_state_rumi_grand_centered, trait_rumination, max_day_esm) %>% 
  distinct() %>% 
  select(-id)

# partial correlation test
pcor(corr_rumi_partial, method = "spearman")

# semi-partial correlation test
spcor(corr_rumi_partial, method= "spearman")
```

## visualization

```{r}
corrplot <- 
  ggplot(corr_rumi, aes(trait_rumination, mean_state_rumi_grand_centered)) +
    geom_point(alpha = 0.5, position = "jitter") +
    # geom_smooth(method = "lm", fill = "#BE7E9D", color = "#9b0a36") +
    geom_smooth(method = "lm", fill = "#C7728F", color = "#8E2438") +
    theme_classic() +
    labs(x = "Trait rumination",
         y = "State rumination\n(aggregated per person)"
    ) +
    theme(axis.title = element_text(size = 12),
    panel.grid.minor.y = element_blank(),
    axis.line = element_line()
    )
corrplot

# ggsave("Figure2.jpg", corrplot, width = 15, height = 10, dpi = 600)

# HU
ggplot(corr_rumi, aes(trait_rumination, mean_state_rumi_grand_centered)) +
  geom_point(alpha = 0.5, position = "jitter") +
  geom_smooth(method = "lm", fill = "#B085BB", color = "#823D93") +
  theme_classic() +
  labs(x = "Vonás-rumináció",
       y = "Állapot-rumináció\n(személyenként átlagolva)"
  ) +
  theme(
    axis.title = element_text(size = 12),
    text = element_text(family = "Cooper Hewitt")
  )
```
