---
title: "06_bootstrapping"
author: "Flora"
date: "2022-12-29"
output: 
  html_document: 
    fig_caption: yes
editor_options: 
  chunk_output_type: inline
---

# loading packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lmeresampler))
suppressPackageStartupMessages(library(boot))

options(scipen = 999)
```

# reading saved model
```{r}
# full_model <- readRDS("saved_models/full_model.rds")
```


# bootstrapping
## residual bootstrap
```{r}
# residual_boot_full_model <-
#   bootstrap(
#     full_model,
#     .f = fixef,
#     type = "residual",
#     B = 5000 ,
#     resample = c(TRUE, TRUE)
#   )
# 
# saveRDS(residual_boot_full_model, "bootstrapping/residual_boot_full_model.rds")
```


### reading bootstrap file and extracting results into a list
```{r}
residual_boot_full_model <- readRDS("bootstrapping/residual_boot_full_model.rds")

list_residual_boot_full_model <-
  list(residual_boot_full_model)
```


### extracting bootstrapped coefficients
```{r paged.print=TRUE}
residual_boot_ci_full_model_basic  <-
  list_residual_boot_full_model %>%
  map(confint, type = "basic")
residual_boot_ci_full_model_basic

residual_boot_ci_full_model_perc  <-
  list_residual_boot_full_model %>%
  map(confint, type = "perc")
residual_boot_ci_full_model_perc
```



## cases bootstrap
```{r}
# cases_boot_full_model <- 
#   bootstrap(
#     full_model,
#     .f = fixef,
#     type = "case",
#     B = 5000 ,
#     resample = c(TRUE, TRUE)
#   )

# saveRDS(cases_boot_full_model, "bootstrapping/cases_boot_full_model.rds")
```


### reading bootstrap file and extracting results into a list
```{r}
# cases_boot_full_model <- readRDS("bootstrapping/cases_boot_full_model.rds")
# 
# list_cases_boot_full_model <- 
#   list(cases_boot_full_model)
```


### extracting bootstrapped coefficients
```{r paged.print=TRUE}
# cases_boot_ci_full_model  <-
#   list_cases_boot_full_model %>%
#   map(confint, type = "norm")
# cases_boot_ci_full_model
```



## parametric bootstrap
```{r}
# parametric_boot_full_model <-
#   bootstrap(
#     full_model,
#     .f = fixef,
#     type = "parametric",
#     B = 5000 ,
#     resample = c(TRUE, TRUE)
#   )
# 
# saveRDS(parametric_boot_full_model, "bootstrapping/parametric_boot_full_model.rds")
```


### reading bootstrap file and extracting results into a list
```{r}
# parametric_boot_full_model <- readRDS("bootstrapping/parametric_boot_full_model.rds")
# 
# list_parametric_boot_full_model <- 
#   list(parametric_boot_full_model)
```


### extracting bootstrapped coefficients
```{r paged.print=TRUE}
# parametric_boot_ci_full_model_basic  <-
#   list_parametric_boot_full_model %>%
#   map(confint, type = "basic")
# parametric_boot_ci_full_model_basic
# 
# parametric_boot_ci_full_model_perc  <-
#   list_parametric_boot_full_model %>%
#   map(confint, type = "perc")
# parametric_boot_ci_full_model_perc
```



# visualizing bootstrap results
## residual bootstrap
### main effects (prediction of negative affect)
```{r echo=TRUE}
# extracting results into a tibble
dat_residual_boot_full_model_maineff <- 
  list_residual_boot_full_model %>% 
  map("replicates") %>% 
  map(
    ~tibble(
      trait_rumination = na.omit(.x$trait_rumination),
      mean_state_rumination = na.omit(.x$mean_state_rumi_grand_centered),
      state_rumination = na.omit(.x$rumination_centered)
    )
  ) %>%
  bind_rows()


# rearranging tibble format
dat_residual_boot_full_model_maineff_long <- 
  dat_residual_boot_full_model_maineff %>% 
  pivot_longer(c(trait_rumination, mean_state_rumination, state_rumination), names_to = "rumination_measure", values_to = "rumination_intensity")


# plotting
coef_residual_boot_full_model_maineff <- 
  dat_residual_boot_full_model_maineff_long %>% 
  ggplot(aes(rumination_intensity, fill = rumination_measure)) +
     geom_vline(xintercept = 0, color = "black", linetype = 2, alpha = .5, size = 1) + 
     geom_density(alpha = .7, color = NA) +
     labs(
       x    = "Boostrapped estimates for the prediction of\nnegative affect by different measures of rumination",
       y = "Density",
       fill = "",
       title    = ""
       ) +
     theme_classic() +
     theme(legend.position = c(0.8, 0.8)) +
     scale_fill_manual(values = c("#FF9500", "#219C19", "#609CCE"), labels = c("Mean state rumination", "State rumination", "Trait rumination"))

coef_residual_boot_full_model_maineff
```


### interactions (moderation of emotional reactivity)
```{r echo=TRUE}
# extracting results into a tibble
dat_residual_boot_full_model_interac <- 
  list_residual_boot_full_model %>% 
  map("replicates") %>% 
  map(
    ~tibble(
      trait_rumination = na.omit(.x$"event_unpleasantness_centered:trait_rumination"),
      mean_state_rumination = na.omit(.x$"event_unpleasantness_centered:mean_state_rumi_grand_centered"),
      state_rumination = na.omit(.x$"event_unpleasantness_centered:rumination_centered")
    )
  ) %>%
  bind_rows()


# rearranging tibble format
dat_residual_boot_full_model_interac_long <- 
  dat_residual_boot_full_model_interac %>% 
  pivot_longer(c(trait_rumination, mean_state_rumination, state_rumination), names_to = "rumination_measure", values_to = "rumination_intensity")


# plotting
coef_residual_boot_full_model_interac <- 
  dat_residual_boot_full_model_interac_long %>% 
  ggplot(aes(rumination_intensity, fill = rumination_measure)) +
     geom_vline(xintercept = 0, color = "black", linetype = 2, alpha = .5, size = 1) + 
     geom_density(alpha = .7, color = NA) +
     labs(
       x    = "Boostrapped estimates for the moderation of\nemotional reactivity by different measures of rumination",
       y = "Density",
       fill = "",
       title    = ""
       ) +
       theme_classic() +
       theme(legend.position = c(0.8, 0.8)) +
       scale_fill_manual(values = c("#FF9500", "#219C19", "#609CCE"), labels = c("Mean state rumination", "State rumination", "Trait rumination"))

coef_residual_boot_full_model_interac
```



## cases bootstrap
### main effects (prediction of negative affect)
```{r echo=TRUE}
# # extracting results into a tibble
# dat_cases_boot_full_model_maineff <- 
#   list_cases_boot_full_model %>% 
#   map("replicates") %>% 
#   map(
#     ~tibble(
#       trait_rumination = na.omit(.x$trait_rumination),
#       mean_state_rumination = na.omit(.x$mean_state_rumi_grand_centered),
#       state_rumination = na.omit(.x$rumination_centered)
#     )
#   ) %>%
#   bind_rows()
# 
# 
# # rearranging tibble format
# dat_cases_boot_full_model_maineff_long <- 
#   dat_cases_boot_full_model_maineff %>% 
#   pivot_longer(c(trait_rumination, mean_state_rumination, state_rumination), names_to = "rumination_measure", values_to = "rumination_intensity")
# 
# 
# # plotting
# coef_cases_boot_full_model_maineff <- 
#   dat_cases_boot_full_model_maineff_long %>% 
#   ggplot(aes(rumination_intensity, fill = rumination_measure)) +
#      geom_vline(xintercept = 0, color = "black", linetype = 2, alpha = .5, size = 1) + 
#      geom_density(alpha = .7, color = NA) +
#      labs(
#        x    = "Boostrapped estimates of the prediction of\nnegative affect by different measures of rumination",
#        y = "Density",
#        fill = "",
#        title    = ""
#        ) +
#      theme_classic() +
#      scale_fill_manual(values = c("#FF9500", "#219C19", "#609CCE"), labels = c("Mean state rumination", "State rumination", "Trait rumination"))
# 
# coef_cases_boot_full_model_maineff
```


### interactions (moderation of emotional reactivity)
```{r echo=TRUE}
# # extracting results into a tibble
# dat_cases_boot_full_model_interac <- 
#   list_cases_boot_full_model %>% 
#   map("replicates") %>% 
#   map(
#     ~tibble(
#       trait_rumination = na.omit(.x$"event_unpleasantness_centered:trait_rumination"),
#       mean_state_rumination = na.omit(.x$"event_unpleasantness_centered:mean_state_rumi_grand_centered"),
#       state_rumination = na.omit(.x$"event_unpleasantness_centered:rumination_centered")
#     )
#   ) %>%
#   bind_rows()
# 
# 
# # rearranging tibble format
# dat_cases_boot_full_model_interac_long <- 
#   dat_cases_boot_full_model_interac %>% 
#   pivot_longer(c(trait_rumination, mean_state_rumination, state_rumination), names_to = "rumination_measure", values_to = "rumination_intensity")
# 
# 
# # plotting
# coef_cases_boot_full_model_interac <- 
#   dat_cases_boot_full_model_interac_long %>% 
#   ggplot(aes(rumination_intensity, fill = rumination_measure)) +
#      geom_vline(xintercept = 0, color = "black", linetype = 2, alpha = .5, size = 1) + 
#      geom_density(alpha = .7, color = NA) +
#      labs(
#        x    = "Boostrapped estimates of the moderation of\nemotional reactivity by different measures of rumination",
#        y = "Density",
#        fill = "",
#        title    = ""
#        ) +
#      theme_classic() +
#      scale_fill_manual(values = c("#FF9500", "#219C19", "#609CCE"), labels = c("Mean state rumination", "State rumination", "Trait rumination"))
# 
# coef_cases_boot_full_model_interac
```



## parametric bootstrap
### main effects (prediction of negative affect)
```{r echo=TRUE}
# # extracting results into a tibble
# dat_parametric_boot_full_model_maineff <- 
#   list_parametric_boot_full_model %>% 
#   map("replicates") %>% 
#   map(
#     ~tibble(
#       trait_rumination = na.omit(.x$trait_rumination),
#       mean_state_rumination = na.omit(.x$mean_state_rumi_grand_centered),
#       state_rumination = na.omit(.x$rumination_centered)
#     )
#   ) %>%
#   bind_rows()
# 
# 
# # rearranging tibble format
# dat_parametric_boot_full_model_maineff_long <- 
#   dat_parametric_boot_full_model_maineff %>% 
#   pivot_longer(c(trait_rumination, mean_state_rumination, state_rumination), names_to = "rumination_measure", values_to = "rumination_intensity")
# 
# 
# # plotting
# coef_parametric_boot_full_model_maineff <- 
#   dat_parametric_boot_full_model_maineff_long %>% 
#   ggplot(aes(rumination_intensity, fill = rumination_measure)) +
#      geom_vline(xintercept = 0, color = "black", linetype = 2, alpha = .5, size = 1) + 
#      geom_density(alpha = .7, color = NA) +
#      labs(
#        x    = "Boostrapped estimate of the prediction of\nnegative affect by different measures of rumination",
#        y = "Density",
#        fill = "",
#        title    = ""
#        ) +
#      theme_classic() +
#      scale_fill_manual(values = c("#FF9500", "#219C19", "#609CCE"), labels = c("Mean state rumination", "State rumination", "Trait rumination"))
# 
# coef_parametric_boot_full_model_maineff
```


### interactions (moderation of emotional reactivity)
```{r echo=TRUE}
# # extracting results into a tibble
# dat_parametric_boot_full_model_interac <- 
#   list_parametric_boot_full_model %>% 
#   map("replicates") %>% 
#   map(
#     ~tibble(
#       trait_rumination = na.omit(.x$"event_unpleasantness_centered:trait_rumination"),
#       mean_state_rumination = na.omit(.x$"event_unpleasantness_centered:mean_state_rumi_grand_centered"),
#       state_rumination = na.omit(.x$"event_unpleasantness_centered:rumination_centered")
#     )
#   ) %>%
#   bind_rows()
# 
# 
# # rearranging tibble format
# dat_parametric_boot_full_model_interac_long <- 
#   dat_parametric_boot_full_model_interac %>% 
#   pivot_longer(c(trait_rumination, mean_state_rumination, state_rumination), names_to = "rumination_measure", values_to = "rumination_intensity")
# 
# 
# # plotting
# coef_parametric_boot_full_model_interac <- 
#   dat_parametric_boot_full_model_interac_long %>% 
#   ggplot(aes(rumination_intensity, fill = rumination_measure)) +
#      geom_vline(xintercept = 0, color = "black", linetype = 2, alpha = .5, size = 1) + 
#      geom_density(alpha = .7, color = NA) +
#      labs(
#        x    = "Boostrapped estimate of the moderation of\nemotional reactivity by different measures of rumination",
#        y = "Density",
#        fill = "",
#        title    = ""
#        ) +
#      theme_classic() +
#      scale_fill_manual(values = c("#FF9500", "#219C19", "#609CCE"), labels = c("Mean state rumination", "State rumination", "Trait rumination"))
# 
# coef_parametric_boot_full_model_interac
```

