---
title: "05_models"
output: html_notebook
---

# About AIC and BIC:

## It seems the "difference of 2" rule of thumb is valid, based on these couple of descriptions I found. So essentially, if a model is 2 AIC units lower than the other, it is considered significantly better.
https://www.scribbr.com/statistics/akaike-information-criterion/
https://onlinelibrary.wiley.com/doi/pdf/10.1002/9781118856406.app5
https://towardsdatascience.com/introduction-to-aic-akaike-information-criterion-9c9ba1c96ced
https://machinelearningmastery.com/probabilistic-model-selection-measures/

## The difference between them (very simply put): AIC is best for prediction, while BIC is best for explanation. Also, BIC penalizes free parameters more strongly.
https://stats.stackexchange.com/questions/577/is-there-any-reason-to-prefer-the-aic-or-bic-over-the-other



# loading packages
```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(sjPlot))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(matrixStats))
suppressPackageStartupMessages(library(performance))
```


# reading dataset
```{r}
dat_model <- read_csv("data/data_model.csv")
```


# baseline model
```{r}
baseline_model <-
  
  lmer(NA_sum ~ age + gender + beep + day + 
      event_unpleasantness_centered +
      (event_unpleasantness_centered | id), data = dat_model)

summary(baseline_model)
tab_model(baseline_model)
AIC(baseline_model)
BIC(baseline_model)
# check_model(baseline_model)
```


# trait rumination
```{r}
trait_model <-
  
  lmer(NA_sum ~ age + gender + beep + day + 
      event_unpleasantness_centered * trait_rumination + 
      (event_unpleasantness_centered | id), data = dat_model)

tab_model(trait_model)
AIC(trait_model)
BIC(trait_model)
# check_model(trait_model)
```


# mean state rumination
```{r}
mean_state <-
  
  lmer(NA_sum ~ age + gender + beep + day + 
      event_unpleasantness_centered * mean_state_rumi_grand_centered + 
      (event_unpleasantness_centered | id), data = dat_model)

tab_model(mean_state)
AIC(mean_state)
BIC(mean_state)
# check_model(mean_state)
```


# state rumination
```{r}
state_model <-
  
  lmer(NA_sum ~ age + gender + beep + day + 
      event_unpleasantness_centered * rumination_centered + 
      (event_unpleasantness_centered * rumination_centered | id), data = dat_model)

tab_model(state_model)
AIC(state_model)
BIC(state_model)
# check_model(state_model)
```
