---
title: "07_models_male"
output:
  html_document:
    df_print: paged
  html_notebook:
    fig_width: 12
    fig_height: 4
editor_options:
  chunk_output_type: console
---

# About AIC and BIC:

## It seems the "difference of 2" rule of thumb is valid, based on these couple of descriptions I found. So essentially, if a model is 2 AIC units lower than the other, it is considered significantly better.
https://www.scribbr.com/statistics/akaike-information-criterion/
https://onlinelibrary.wiley.com/doi/pdf/10.1002/9781118856406.app5
https://towardsdatascience.com/introduction-to-aic-akaike-information-criterion-9c9ba1c96ced
https://machinelearningmastery.com/probabilistic-model-selection-measures/

## The difference between them (very simply put): AIC is best for prediction, while BIC is best for explanation. Also, BIC penalizes free parameters more strongly.
https://stats.stackexchange.com/questions/577/is-there-any-reason-to-prefer-the-aic-or-bic-over-the-other



# loading packages
```{r}
suppressPackageStartupMessages(library(sjPlot))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(library(matrixStats))
suppressPackageStartupMessages(library(performance))
suppressPackageStartupMessages(library(see))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(nortest))
suppressPackageStartupMessages(library(extrafont))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(showtext))
suppressPackageStartupMessages(library(lmeresampler))
suppressPackageStartupMessages(library(boot))
suppressPackageStartupMessages(library(tidyverse))

loadfonts(device = "all")
font_add(family = "Cooper Hewitt", regular = "CooperHewitt-Medium.otf")
showtext_auto()

options(scipen = 999)
```


# reading dataset
```{r message=FALSE}
dat_model_male <- read_csv("data/data_model.csv") %>% 
  filter(gender == "Male")
```


# baseline model
```{r paged.print=FALSE}
# baseline_model <- 
#   
#   lmer(NA_sum ~ age + gender + beep + day + 
#       event_unpleasantness_centered +
#       (event_unpleasantness_centered | id), data = dat_model_male)
# 
# summary(baseline_model)
# tab_model(baseline_model, show.std = T, digits = 3, digits.re = 3)
# AIC(baseline_model)
# BIC(baseline_model)
# check_collinearity(baseline_model)
# # check_autocorrelation(baseline_model)
# # check_model(baseline_model)
# plot_model(baseline_model, type = "diag")
# plot_model(baseline_model, type = "eff")
# 
# # normality test for residuals
# baseline_model_residuals <- residuals(baseline_model)
# lillie.test(baseline_model_residuals)
# 
# # normality test for NA_sum
# lillie.test(dat_model_male$NA_sum)
```


# baseline model with log10-transformed NA
```{r paged.print=FALSE}
# baseline_model_NA_log <-
#   
#   lmer(NA_sum_log ~ age + gender + beep + day + 
#       event_unpleasantness_centered +
#       (event_unpleasantness_centered | id), data = dat_model_male)
# 
# summary(baseline_model_NA_log)
# tab_model(baseline_model_NA_log, show.std = T)
# AIC(baseline_model_NA_log)
# BIC(baseline_model_NA_log)
# # check_model(baseline_model_NA_log)
# plot_model(baseline_model_NA_log, type = "diag")
# 
# 
# # normality test for residuals
# baseline_model_NA_log_residuals <- residuals(baseline_model_NA_log)
# lillie.test(baseline_model_NA_log_residuals)
```


# trait rumination
```{r}
trait_model_male <-

  lmer(NA_sum ~ age + beep + day +
      trait_rumination +
      (1 | id), data = dat_model_male)

tab_model(trait_model_male, show.std = T)
AIC(trait_model_male)
BIC(trait_model_male)
# check_model(trait_model_male)
```


## coefplot for trait rumination
```{r}
coefplot_trait <- 
  plot_model(trait_model,
             type = "std",
             rm.terms = c("age", "gender [Male]", "gender [Other]", "beep", "day", "event_unpleasantness_centered"),
             ci.lvl = 0.95,
             std.est = T,
             title = "",
             show.values = TRUE,
             colors = "#7581A7",
             value.size = 6,
             dot.size = 3,
             line.size = 1.2,
             axis.labels = c("Trait rumination * Perceived stress", "Trait rumination"),
             axis.title = "Standardized β coefficients"
             ) +
             ylim(0, 0.6) +
             theme_minimal() +
             theme(axis.text = element_text(size = 14, color = "black"),
                   axis.title.x = element_text(size = 14, color = "black"),
                   axis.line = element_line(size = .3),
                   # text = element_text(family = "Cooper Hewitt")
             )
coefplot_trait
```


# mean state rumination
```{r}
mean_state_model_male <-

  lmer(NA_sum ~ age + beep + day +
      mean_state_rumi_grand_centered +
      (1 | id), data = dat_model_male)

tab_model(mean_state_model_male, show.std = T)
AIC(mean_state_model_male)
BIC(mean_state_model_male)
# check_model(mean_state_model_male)
```


## coefplot for mean state rumination
```{r}
coefplot_mean_state <- 
  plot_model(mean_state_model,
             type = "std",
             rm.terms = c("age", "gender [Male]", "gender [Other]", "beep", "day", "event_unpleasantness_centered"),
             ci.lvl = 0.95,
             std.est = T,
             title = "",
             show.values = TRUE,
             colors = "#FFBF69",
             value.size = 6,
             dot.size = 3,
             line.size = 1.2,
             axis.labels = c("Mean state rumination * Perceived stress", "Mean state rumination"),
             axis.title = "Standardized β coefficients"
             ) +
             ylim(0, 0.6) +
             theme_minimal() +
             theme(axis.text = element_text(size = 14, color = "black"),
                   axis.title.x = element_text(size = 14, color = "black"),
                   axis.line = element_line(size = .3),
                   # text = element_text(family = "Cooper Hewitt")
             )
coefplot_mean_state
```


# state rumination
```{r}
state_model_male <-

  lmer(NA_sum ~ age + beep + day +
      rumination_centered +
      (1 | id), data = dat_model_male)

tab_model(state_model_male, show.std = T)
tab_model(trait_model_male, mean_state_model_male, state_model_male, show.std = T, show.aic = T)
AIC(state_model_male)
BIC(state_model_male)
# check_model(state_model_male)
```


## coefplot for state rumination
```{r}
coefplot_state <- 
  plot_model(state_model,
             type = "std",
             rm.terms = c("age", "gender [Male]", "gender [Other]", "beep", "day", "event_unpleasantness_centered"),
             ci.lvl = 0.95,
             std.est = T,
             title = "",
             show.values = TRUE,
             colors = "#B0CDB6",
             value.size = 6,
             dot.size = 3,
             line.size = 1.2,
             axis.labels = c("State rumination * Perceived stress", "State rumination"),
             axis.title = "Standardized β coefficients"
             ) +
             ylim(0, 0.6) +
             theme_minimal() +
             theme(axis.text = element_text(size = 14, color = "black"),
                   axis.title.x = element_text(size = 14, color = "black"),
                   axis.line = element_line(size = .3),
                   # text = element_text(family = "Cooper Hewitt")
             )
coefplot_state
```


# full model negaff
```{r}
full_model_male_negaff <-
  
  lmer(NA_sum ~
      rumination_centered +  
      mean_state_rumi_grand_centered +
      trait_rumination +
      
      (rumination_centered | id), data = dat_model_male,
      control = lmerControl(optCtrl=list(ftol_abs=1e-8,xtol_abs=1e-8))
  )

tab_model(full_model_male_negaff, show.std = T, digits = 3, digits.re = 3)
summary(full_model_male_negaff)
plot_model(full_model_male_negaff, type = "eff")
# check_model(full_model_male_negaff)
check_collinearity(full_model_male_negaff)
plot_model(full_model_male_negaff, rm.terms = c("age", "gender [Male]", "gender [Other]", "beep", "day", "event_unpleasantness_centered"), ci.lvl=0.95)
AIC(full_model_male_negaff)
BIC(full_model_male_negaff)

# checking assumptions
plot_model(full_model_male_negaff)
```

# full model posaff
```{r}
full_model_male_posaff <-
  
  lmer(PA_sum ~
      rumination_centered +  
      mean_state_rumi_grand_centered +
      trait_rumination +
      
      (rumination_centered | id), data = dat_model_male,
      control = lmerControl(optCtrl=list(ftol_abs=1e-8,xtol_abs=1e-8))
  )

tab_model(full_model_male_posaff, show.std = T, digits = 3, digits.re = 3)
summary(full_model_male_posaff)
plot_model(full_model_male_posaff, type = "eff")
# check_model(full_model_male_posaff)
check_collinearity(full_model_male_posaff)
plot_model(full_model_male_posaff, rm.terms = c("age", "gender [Male]", "gender [Other]", "beep", "day", "event_unpleasantness_centered"), ci.lvl=0.95)
AIC(full_model_male_posaff)
BIC(full_model_male_posaff)

# checking assumptions
plot_model(full_model_male_posaff)
```

# coefplots for full model negaff
```{r}
# plot for interactions (colored afterwards in photoshop)
coefplot_interactions <- 
  plot_model(full_model,
             type = "std",
             rm.terms = c("rumination_centered", "trait_rumination", "mean_state_rumi_grand_centered", "age", "gender [Male]", "gender [Other]",   "beep", "day", "event_unpleasantness_centered"),
             ci.lvl = 0.95,
             std.est = T,
             title = "",
             show.values = TRUE,
             colors = "black",
             value.size = 6,
             value.offset = .25,
             dot.size = 4,
             line.size = 1.2,
             axis.labels = c("Trait rumination * Perceived stress", "Mean state rumination * Perceived stress", "State rumination * Perceived   stress"),
             axis.title = "Standardized β coefficients"
             ) +
             ylim(0, 0.1) +
             theme_minimal() +
             theme(axis.text = element_text(size = 10, color = "black"),
                   axis.title.x = element_text(size = 12, color = "black"),
                   axis.line = element_line(size = .3),
                   text = element_text(family = "Cooper Hewitt")
             )
coefplot_interactions

# plot for main effects (colored afterwards in photoshop)
coefplot_maineffects <- 
  plot_model(full_model,
             type = "std",
             terms = c("rumination_centered", "mean_state_rumi_grand_centered", "trait_rumination"),
             ci.lvl = 0.95,
             std.est = TRUE,
             title = "",
             show.values = TRUE,
             colors = "black",
             value.size = 6,
             value.offset = 0.25,
             dot.size = 4,
             line.size = 1.2,
             axis.labels = c("Trait rumination", "Mean state rumination", "State rumination"),
             axis.title = "Standardized β coefficients"
             ) +
             ylim(0, 1) +
             theme_minimal() +
             theme(axis.text = element_text(size = 10, color = "black"),
                   axis.title.x = element_text(size = 12, color = "black"),
                   axis.line = element_line(size = .3),
                   text = element_text(family = "Cooper Hewitt")
             )
coefplot_maineffects
```


# coefplots for full model [HU]
```{r}
# plot for interactions (colored afterwards in photoshop)
plot_model(full_model,
           type = "std",
           rm.terms = c("rumination_centered", "trait_rumination", "mean_state_rumi_grand_centered", "age", "gender [Male]", "gender [Other]", "beep", "day", "event_unpleasantness_centered"),
           ci.lvl = 0.95,
           std.est = T,
           title = "",
           show.values = TRUE,
           colors = "black",
           value.size = 4,
           dot.size = 3,
           line.size = 1.2,
           axis.labels = c("vonás-rumináció * észlelt stressz", "átlagos állapot-rumináció * észlelt stressz", "állapot-rumináció * észlelt stressz"),
           axis.title = "standardizált béta együtthatók"
           ) +
           ylim(0, 0.1) +
           theme_minimal() +
           theme(axis.text = element_text(size = 10, color = "black"),
                 axis.title.x = element_text(size = 12, color = "black"),
                 axis.line = element_line(size = .3),
                 text = element_text(family = "Cooper Hewitt")
           )

# plot for main effects (colored afterwards in photoshop)
plot_model(full_model,
           type = "std",
           terms = c("rumination_centered", "mean_state_rumi_grand_centered", "trait_rumination"),
           ci.lvl = 0.95,
           std.est = TRUE,
           title = "",
           show.values = TRUE,
           colors = "black",
           value.size = 4,
           dot.size = 3,
           line.size = 1.2,
           axis.labels = c("vonás-rumináció", "átlagos állapot-rumináció", "állapot-rumináció"),
           axis.title = "standardizált béta együtthatók"
           ) +
           ylim(0, 1) +
           theme_minimal() +
           theme(axis.text = element_text(size = 10, color = "black"),
                 axis.title.x = element_text(size = 12, color = "black"),
                 axis.line = element_line(size = .3),
                 text = element_text(family = "Cooper Hewitt")
           )
```


# full model with log10-transformed NA
```{r}
# full_model_NA_log <-
#   
#   lmer(NA_sum_log ~ age + gender + beep + day + 
#       event_unpleasantness_centered * rumination_centered + 
#       event_unpleasantness_centered * trait_rumination +
#       event_unpleasantness_centered * mean_state_rumi_grand_centered +
#       (event_unpleasantness_centered * rumination_centered | id), data = dat_model_male
#       # control = lmerControl(optCtrl=list(ftol_abs=1e-8,xtol_abs=1e-8))
#   )
# 
# tab_model(full_model_NA_log, show.std = T)
# summary(full_model_NA_log)
# plot_model(full_model_NA_log, type = "diag")
# AIC(full_model_NA_log)
# BIC(full_model_NA_log)
```


# saving models
```{r}
# saveRDS(full_model, "saved_models/full_model.rds")
```


# bootstrapping full model negaff
## running bootstrap
```{r}
case_boot_full_model_male_negaff <-
  bootstrap(
    full_model_male_negaff,
    .f = fixef,
    type = "case",
    B = 5000 ,
    resample = c(TRUE, TRUE)
  )

saveRDS(case_boot_full_model_male_negaff, "bootstrapping/case_boot_full_model_male_negaff.rds")
```

## reading bootstrap file and extracting results into a list
```{r}
case_boot_full_model_male_negaff <- readRDS("bootstrapping/case_boot_full_model_male_negaff.rds")

list_case_boot_full_model_male_negaff <-
  list(case_boot_full_model_male_negaff)
```

## extracting bootstrapped coefficients
```{r paged.print=TRUE}
case_boot_ci_full_model_male_negaff_basic  <-
  list_case_boot_full_model_male_negaff %>%
  map(confint, type = "basic")
case_boot_ci_full_model_male_negaff_basic

case_boot_ci_full_model_male_negaff_perc  <-
  list_case_boot_full_model_male_negaff %>%
  map(confint, type = "perc")
case_boot_ci_full_model_male_negaff_perc
```


# bootstrapping full model posaff
## running bootstrap
```{r}
case_boot_full_model_male_posaff <-
  bootstrap(
    full_model_male_posaff,
    .f = fixef,
    type = "case",
    B = 5000 ,
    resample = c(TRUE, TRUE)
  )

saveRDS(case_boot_full_model_male_posaff, "bootstrapping/case_boot_full_model_male_posaff.rds")
```

## reading bootstrap file and extracting results into a list
```{r}
case_boot_full_model_male_posaff <- readRDS("bootstrapping/case_boot_full_model_male_posaff.rds")

list_case_boot_full_model_male_posaff <-
  list(case_boot_full_model_male_posaff)
```

## extracting bootstrapped coefficients
```{r paged.print=TRUE}
case_boot_ci_full_model_male_posaff_basic  <-
  list_case_boot_full_model_male_posaff %>%
  map(confint, type = "basic")
case_boot_ci_full_model_male_posaff_basic

case_boot_ci_full_model_male_posaff_perc  <-
  list_case_boot_full_model_male_posaff %>%
  map(confint, type = "perc")
case_boot_ci_full_model_male_posaff_perc
```
